Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-12 appendix.tex
updating results

\subsubsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
The real and imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag
  =\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \\
   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
  =\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right],
\end{align}
where we recall $\hat{\omega} = \real \hat{\sigma}$ and
$\hat{\nu}=\imag\hat{\sigma}$. 
Adding the square of these equations give
\begin{align}
  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}
  =\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime
  \right|^2d\hat{z} 
  \leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  \left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z}
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
vertical shear in the domain considered.

@@

\subsection{Explicit solutions in the thin-disk limit}\label{iso_explicit}
Here we assume a thin disk ($\epsilon\ll1$) so that $\ln\rho\simeq
-\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume
low-frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \frac{\ii q\epsilon \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
\end{align}
We remark that for $\hat{k}^2\gg 1$ and taking low-frequency limit,
Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in \cite{nelson13},
although we have taken a different route.

@@

\subsubsection{Stability without vertical shear}
In the absence of vertical shear $q=0$, Eq. \ref{simple_growth} can be
written as \begin{align}
  \hat{\sigma}^2\khat^2 =
  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). 
\end{align}
This is just the dispersion relation for axisymmetric waves in a
vertically isothermal disk already discussed in previous works 
\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In
this case the solutions are Hermite polynomials, $W\propto
\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
is stable, consistent with \S\ref{integral_relation}.

>>>>

\subsubsection{Maximum growth rate}\label{max_growth1}
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
The real and imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag
  =\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \\
   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
  =\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right],
\end{align}
where we recall $\hat{\omega} = \real \hat{\sigma}$ and
$\hat{\nu}=\imag\hat{\sigma}$. 
Adding the square of these equations give
\begin{align}
  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}
  =\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime
  \right|^2d\hat{z} 
  \leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  \left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z}
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
vertical shear in the domain considered.

@@

\subsection{Explicit solutions in the thin-disk limit}\label{iso_explicit}
Here we assume a thin disk ($\epsilon\ll1$) so that $\ln\rho\simeq
-\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume 
low-frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \frac{\ii q\epsilon \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
\end{align}
We remark that taking the low-frequency limit and considering
$\hat{k}^2\gg 1$, Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in
\cite{nelson13}, although we have taken a different route.

@@

\subsubsection{Stability without vertical shear}
In the absence of vertical shear $q=0$, Eq. \ref{simple_growth} can be
written as \begin{align}
  \hat{\sigma}^2\khat^2 =
  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). \label{iso_disp_full}
\end{align}
This is just the dispersion relation for axisymmetric waves in a
vertically isothermal disk already discussed in previous works 
\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In
this case the solutions are Hermite polynomials, $W\propto
\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
is stable, consistent with \S\ref{integral_relation}. The
low-frequency solution to Eq. \ref{iso_disp_full} correspond to
inertial waves \citep{balbus03}.
